---
title: "lab1-gfan"
author: "gucci fan"
date: "2024-10-24"
output: html_document
---

I wanted to find a dataset similar to the example Boulder data but for a different city and perform analysis on that. I liked the Boulder elevation maps so I wanted to recreate that from publicly available data. I took elevation data for a section of Phoenix, AZ from USGS but it was an extremely large and detailed dataset that was too big to load (420 mb) into a dataframe. So I decided to simplify the dataset by reprojecting it in QGIS to have a lower resolution.

I also originally classified the elevation raster into 4 classes: 0-500ft, 500-1000ft, 1000-1500ft, and 1500-2400ft. I exported it as a .tif and visualized it in the RMD.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(sf)
library(tidyverse)
library(raster)
library(viridis) 
```

```{r}
elevation_raster <- raster("test.tif")

elevation_df <- as.data.frame(as(elevation_raster, "SpatialPixelsDataFrame"))
colnames(elevation_df) <- c("elevation", "x", "y")


ggplot(elevation_df) +
  geom_tile(aes(x = x, y = y, fill = elevation)) +
  scale_fill_viridis_c(option = "magma", name = "Elevation Class") +
  theme_minimal() +
  labs(title = "Phoenix Elevation")


```
However this is kind of ugly and undetailed so I went with my original simplified version of the USGS dataset, which both reduces the complexity of the original 420mb dataset while keeping more of the details than what I had in the 4-class map - this one will interpret it as a continuous range of colors rather than discrete values. Since it's elevation, I think continuous colors are better anyway.

```{r}

og_elev_raster <- raster("originalelev.tif")
og_elev_df <- as.data.frame(as(og_elev_raster, "SpatialPixelsDataFrame"))
colnames(og_elev_df) <- c("elevation", "Longitude", "Latitude")

og_elev_df$elevation_class_quantile <- cut(og_elev_df$elevation, breaks = quantile(og_elev_df$elevation, probs = seq(0, 1, by = 0.25)), include.lowest = TRUE)

ggplot(og_elev_df) +
  geom_tile(aes(x = Longitude, y = Latitude, fill = elevation)) +
  scale_fill_viridis_c(option = "viridis", name = "Elevation (ft)") +
  theme_minimal() +
  labs(title = "Phoenix Elevation")
```
Lastly, assuming the 40,000 data points in my elevation dataset are taken at regularly spaced intervals, I thought it would be interesting to display the spread and frequency of elevation values in a histogram. From observing the map it appears the vast majority of the sampled area is in the lower 50% of elevation values, with some peaks at the northeast corner. When constructing the histogram, the result follows that pattern, with a very right-skewed distribution favoring lower elevation values.

```{r}
# Using ggplot2 to create a histogram
ggplot(og_elev_df, aes(x = elevation)) +
  geom_histogram(bins = 20 , fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Phoenix Elevation Values", x = "Elevation (ft)", y = "Frequency")

```

